Data exploration, quality control and statistical analysis of ChIP-exo experiments

### 2016-04-28 rmarkdown/paper.tex
added outline for paper

\begin{table}[h!]
%% need to fix format
  \centering

@@

\subsection*{Evaluation of QC visualizations}

{\color{blue} \emph{need to describe figure \ref{fig:enrich_all}
}
\color{black}

@@

We considered dPeak Chung et al., 2013 \cite{dpeak}, GEM Guo et al.,
2012 \cite{gem}, MACE Wang et al., 2014 \cite{mace} and Peazilla
Bardet et al., 2013 \cite{peakzilla} for the ChIP-exo data
analysis. for the dPeak algorithm we used the R package \textbf{dPeak}
version 2.0.1 which is available from
\url{https://github.com/dongjunchung/dpeak}. For the GEM algorithm, we
used it's Java implementation version 2.6 which is available from
\url{http://groups.csail.mit.edu/cgs/gem/}. For the Mace algorithm, we
used it Python implementation version 1.2, which is available from
\url{http://dldcc-web.brc.bcm.edu/lilab/MACE/docs/html/}. For the
Peakzilla algorithm, we used the version available in
\url{https://github.com/steinmann/peakzilla}. Candidate regions for
\textbf{dPeak} were identified for each replicate of ChIP-exo data
using the \textbf{MOSAiCS} algorithm Kuan et al., 2011 \cite{mosaics}
(one sample analysis using false discovery rate of 0.01\%)
implemented as an R package \textbf{mosaics} version 2.9.7 (available
from \emph{bioconductor}). We further filtered out candidate regions
by using the 300 peaks with higher average ChIP tag count to
avoid potential false positive based on the exploratory
analysis. These regions were also explicitly provided to the GEM
algorithm as candidate regions. Default tuning parameters were used
during model fitting for all methods. We were unable to use CexoR
\cite{cexor} to estimate ChIP-exo binding sites.

\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{chip_exo_paper}
\nocite{exo_gb}
\nocite{maplot1}
\nocite{maplot2}
\nocite{chipbeyond}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width = .9\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
  \caption{Comparison of the number of candidate regions (A),
    predicted events (B), identified targets (C) and resolution (D)
    among ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations
    are considered as a gold standard. A gold standard binding events
    was deemed identified if a binding event was estimated at a $\pm$
    15 vicinity of it.}
  \label{fig:design}
\end{figure}

@@

\begin{table}[h!]
%% need to fix format
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{verbatim}
##    Organism IP/TF Condition/Cell Rep.      Depth    PBC     NSC Protocol
## 1:   E.Coli Sig70       Rif-0min    1 13,445,022 0.9426  8.8652       PE
## 2:   E.Coli Sig70       Rif-0min    1 16,538,920 0.9379  7.0318       PE
## 3:   E.Coli Sig70      Rif-20min    2 11,642,722 0.8892 10.7728       PE
## 4:   E.Coli Sig70      Rif-20min    2 16,854,026 0.9407  7.9362       PE
## 5:   E.Coli Sig70       Rif-0min    1  6,722,511 0.6633  9.0178       SE
## 6:   E.Coli Sig70       Rif-0min    1  8,269,460 0.5594  7.1795       SE
## 7:   E.Coli Sig70      Rif-20min    2  5,821,361 0.6472 10.8990       SE
## 8:   E.Coli Sig70      Rif-20min    2  8,427,013 0.5895  8.1247       SE
\end{verbatim}
\end{kframe}
\end{knitrout}
\caption{Same QC metrics as in table \ref{tab:qc} but applied to
  Landick's chipseq data of the rif experiment}
\end{table}

>>>>

\begin{table}[h!]
%% need to fix format
  \centering
%% need to re-build this table

@@

% \subsection*{Evaluation of QC visualizations}

% Each boxplot in Figure \ref{fig:enrich_all}A, was created by sampling
% Msamp of the regions generated and fitting the model:

% \begin{align*}
%   \mbox{depth} = \beta \times \text{(\# unique
%       positions)}
% \end{align*}

% and repeating the process Ntimes times.

%     Intercept and C) Slope in the the $\mbox{URCR} = k / \mbox{ARC} +
%     b$ model f

@@

We considered dPeak Chung et al., 2013 \cite{dpeak}, GEM Guo et al.,
2012 \cite{gem}, MACE Wang et al., 2014 \cite{mace} and Peazilla
Bardet et al., 2013 \cite{peakzilla} for the ChIP-exo data
analysis. for the dPeak algorithm we used the R package \textbf{dPeak}
version 2.0.1 which is available from
\url{https://github.com/dongjunchung/dpeak}. For the GEM algorithm, we
used it's Java implementation version 2.6 which is available from
\url{http://groups.csail.mit.edu/cgs/gem/}. For the Mace algorithm, we
used it Python implementation version 1.2, which is available from
\url{http://dldcc-web.brc.bcm.edu/lilab/MACE/docs/html/}. For the
Peakzilla algorithm, we used the version available in
\url{https://github.com/steinmann/peakzilla}. Candidate regions for
\textbf{dPeak} were identified for each replicate of ChIP-exo data
using the \textbf{MOSAiCS} algorithm Kuan et al., 2011 \cite{mosaics}
(one sample analysis using false discovery rate of 0.01\%)
implemented as an R package \textbf{mosaics} version 2.9.7 (available
from \emph{bioconductor}). We further filtered out candidate regions
by using the 300 peaks with higher average ChIP tag count to
avoid potential false positive based on the exploratory
analysis. These regions were also explicitly provided to the GEM
algorithm as candidate regions. Default tuning parameters were used
during model fitting for all methods. We downloaded \textbf{CexoR}
\cite{cexor} version on its 1.8 version from \emph{bioconductor} but
were unable to use it for the $\sig$ experiments.

 % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliographystyle{vancouver}
\bibliography{chip_exo_paper}
\nocite{exo_gb}
\nocite{maplot1}
\nocite{maplot2}
\nocite{chipbeyond}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width = .9\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
  \caption{Comparison of the number of A) candidate regions, B)
    predicted events, C) identified targets and D) resolution among
    ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are
    considered as a gold standard. A gold standard binding events was
    deemed identified if a binding event was estimated at a $\pm$
    15 vicinity of it.}
  \label{fig:design}
\end{figure}

@@

\begin{table}[h!]
%% need to fix format
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{verbatim}
##    Organism IP/TF Condition/Cell Rep.      Depth    PBC     NSC Protocol
## 1:   E.Coli Sig70       Rif-0min    1 13,445,022 0.9426  8.8652       PE
## 2:   E.Coli Sig70       Rif-0min    1 16,538,920 0.9379  7.0318       PE
## 3:   E.Coli Sig70      Rif-20min    2 11,642,722 0.8892 10.7728       PE
## 4:   E.Coli Sig70      Rif-20min    2 16,854,026 0.9407  7.9362       PE
## 5:   E.Coli Sig70       Rif-0min    1  6,722,511 0.6633  9.0178       SE
## 6:   E.Coli Sig70       Rif-0min    1  8,269,460 0.5594  7.1795       SE
## 7:   E.Coli Sig70      Rif-20min    2  5,821,361 0.6472 10.8990       SE
## 8:   E.Coli Sig70      Rif-20min    2  8,427,013 0.5895  8.1247       SE
\end{verbatim}
\end{kframe}
\end{knitrout}
\caption{Usual ChIP-Seq QC metrics as in table \ref{tab:qc} but
  applied to Landick's chipseq data of the rif experiment}
\end{table}
